Antigenic phenotypes and phylogenetic metrics improve forecasts of seasonal influenza A/H3N2 evolution

### 2019-10-28 manuscript/main.tex
Added missing delta time parameter in forecasting formula
Thanks to Tal Einav for catching this error.
Add cross-validation schemes for simulated and natural data
Note that the cross-validation figure for natural data still only reflects
validation timepoints, as we have not yet analyzed the test data for these
populations.

The long-term evolution of influenza A/H3N2 hemagglutinin has been previously described as a balance between positive selection for substitutions that enable escape from adaptive immunity by modifying existing epitopes and purifying selection on domains that are required to maintain the protein's primary functions of binding and membrane fusion \cite{Bush:1999vj,Neher2013,Luksza:2014hj,Koelle:2015dh}.
To test the ability of our models to accurately detect these evolutionary patterns under controlled conditions, we simulated the long-term evolution of A/H3N2-like viruses under positive and purifying selection for 40 years (Methods).
These selective constraints produced phylogenetic structures and accumulation of epitope and non-epitope mutations that were consistent with phylogenies of natural A/H3N2 HA (Supplemental Figure \ref{sup_fig:simulated_h3n2_ha_phylogeny}).
We fit models to these simulated populations using all sequence-only fitness metrics.

@@

Next, we trained and validated models for individual fitness predictors using 25 years of natural A/H3N2 populations spanning from October 1, 1990 to October 1, 2015.
We held out samples collected between October 1, 2015 and October 1, 2019 for model testing.
In addition to the sequence-only models we tested on simulated populations, we also fit models for our new fitness metrics based on experimental phenotypes including HI cross-immunity and DMS mutational effects.
We hypothesized that both HI and DMS metrics would be assigned positive coefficients, as they estimate increased antigenic drift and beneficial mutations, respectively.
As antigenic drift is generally considered to be the primary evolutionary pressure on natural A/H3N2 populations \cite{Smith:2004jc,Bedford:2014bf,Luksza:2014hj}, we expected that epitope and HI cross-immunity would be individually more predictive than non-epitope mutations or DMS mutational effects.
Previous research \cite{Neher:2016hy} and our simulation results also led us to expect that LBI and delta frequency would outperform other individual mechanistic metrics.

@@

\begin{equation}
    \mathbf{\hat{x}}(t + \Delta{t}) = \frac{1}{Z(t)}\sum_{i}x_{i}(t)\exp(f_{i}(t))
    \label{equation_exponential_growth_model}
\end{equation}

>>>>

The long-term evolution of influenza A/H3N2 hemagglutinin has been previously described as a balance between positive selection for substitutions that enable escape from adaptive immunity by modifying existing epitopes and purifying selection on domains that are required to maintain the protein's primary functions of binding and membrane fusion \cite{Bush:1999vj,Neher2013,Luksza:2014hj,Koelle:2015dh}.
To test the ability of our models to accurately detect these evolutionary patterns under controlled conditions, we simulated the long-term evolution of A/H3N2-like viruses under positive and purifying selection for 40 years (Methods, Supplemental Figure~\ref{sup_fig:cross_validation_for_simulated_populations}).
These selective constraints produced phylogenetic structures and accumulation of epitope and non-epitope mutations that were consistent with phylogenies of natural A/H3N2 HA (Supplemental Figure \ref{sup_fig:simulated_h3n2_ha_phylogeny}).
We fit models to these simulated populations using all sequence-only fitness metrics.

@@

Next, we trained and validated models for individual fitness predictors using 25 years of natural A/H3N2 populations spanning from October 1, 1990 to October 1, 2015.
We held out samples collected between October 1, 2015 and October 1, 2019 for model testing (Supplemental Figure~\ref{sup_fig:cross_validation_for_natural_populations}).
In addition to the sequence-only models we tested on simulated populations, we also fit models for our new fitness metrics based on experimental phenotypes including HI cross-immunity and DMS mutational effects.
We hypothesized that both HI and DMS metrics would be assigned positive coefficients, as they estimate increased antigenic drift and beneficial mutations, respectively.
As antigenic drift is generally considered to be the primary evolutionary pressure on natural A/H3N2 populations \cite{Smith:2004jc,Bedford:2014bf,Luksza:2014hj}, we expected that epitope and HI cross-immunity would be individually more predictive than non-epitope mutations or DMS mutational effects.
Previous research \cite{Neher:2016hy} and our simulation results also led us to expect that LBI and delta frequency would outperform other individual mechanistic metrics.

@@

\begin{equation}
    \mathbf{\hat{x}}(t + \Delta{t}) = \frac{1}{Z(t)}\sum_{i}x_{i}(t)\exp(f_{i}(t)\Delta{t})
    \label{equation_exponential_growth_model}
\end{equation}

### 2019-10-28 manuscript/supplement.tex
Add cross-validation schemes for simulated and natural data
Note that the cross-validation figure for natural data still only reflects
validation timepoints, as we have not yet analyzed the test data for these
populations.

@@

>>>>

\begin{figure*}[h]
  \begin{center}
  \includegraphics[width=\textwidth]{figures/cross-validation-for-simulated-populations.png}
  \caption{
  Time-series cross-validation scheme for simulated populations.
  Models were trained in six-year sliding windows (grey lines) and validated on out-of-sample data from validation timepoints (filled circles).
  Validation results from 30 years of data were used to iteratively tune model hyperparameters.
  After fixing hyperparameters, model coefficients were fixed at the mean values across all training windows.
  Fixed coefficients were applied to 10 years of new out-of-sample test data (open circles) to estimate true forecast errors.
  }
  \label{sup_fig:cross_validation_for_simulated_populations}
  \end{center}
\end{figure*}

@@

\begin{figure*}[h]
  \begin{center}
  \includegraphics[width=\textwidth]{figures/cross-validation-for-natural-populations.png}
  \caption{
  Time-series cross-validation scheme for natural populations.
  Models were trained in six-year sliding windows (grey lines) and validated on out-of-sample data from validation timepoints (filled circles).
  Validation results from 25 years of data were used to iteratively tune model hyperparameters.
  After fixing hyperparameters, model coefficients were fixed at the mean values across all training windows.
  Fixed coefficients were applied to four years of new out-of-sample test data (open circles) to estimate true forecast errors.
  }
  \label{sup_fig:cross_validation_for_natural_populations}
  \end{center}
\end{figure*}
